---
title: "Competing risk anslysis-2"
author: "Nancy"
date: "May 2, 2018"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    theme: united
    highlight: tango
    fig_width: 5
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse,data.table,gridExtra,tableone,survival,magrittr)


ramq_cc<-readRDS('ramq_cc_as_death2.RData')
ramq_cc$status<-factor(ramq_cc$status,levels=c(0,1),labels=c(0,1))
```
#### Case Definition:
Cases were defined as a composite of 

1.having an AS diagnostic code or a surgical valve replacement after cohort entry

2.death of all cause between cohort entry date and 2011-03-31

This analysis aimed to check competing risk, taking into account of death of all cause.Since the cause of death was not accurately documented in the death registry, this sensitivity analysis aimed to assess the effect of missed cases from death due to AS. 


#### Check matching algorithm {.tabset}
##### Distribution of length of follow-up between cases and controls {-}
```{r,fig.height=4,fig.width=6}
ggplot(data=subset(ramq_cc,tfu<1000),aes(x=tfu,fill=status))+
    geom_histogram(binwidth=30,alpha=0.4)
```

```{r}
tapply(ramq_cc$tfu,ramq_cc$status,summary)
```

##### Distribution of index age between cases and controls:{-}
```{r,fig.height=4,fig.width=6}
ggplot(data=ramq_cc,aes(x=tageind,fill=status))+
    geom_histogram(alpha=0.3)

tapply(ramq_cc$tageind,ramq_cc$status,summary)
```

##### Distribution of Year of entry and exit between cases and controls: {-}
```{r,fig.height=4,fig.width=6}
ggplot(data=ramq_cc,aes(x=year(tentry_date),fill=status))+
    geom_bar(alpha=0.3)+scale_x_continuous(breaks=unique(year(ramq_cc$tentry_date)))+
    ggtitle('Year of Entry')+
    scale_fill_discrete(name='Outcome',labels=c('Control','Case'))

ggplot(data=ramq_cc,aes(x=year(texit_date),fill=status))+
    geom_bar(alpha=0.3)+scale_x_continuous(breaks=unique(year(ramq_cc$texit_date)))+
    ggtitle('Year of Exit')+
    scale_fill_discrete(name='Outcome',labels=c('Control','Case'))

ggplot(data=subset(ramq_cc,status==1),aes(x=year(texit_date)))+
    geom_bar(alpha=0.3)+scale_x_continuous(breaks=unique(year(ramq_cc$texit_date)))+
    ggtitle('Year of Exit for Cases')
    

```

#### Percent of cases sampled as controls:
```{r}
control_to_case<-round(100*sum(ramq_cc$tcase==1&ramq_cc$status==0)/sum(ramq_cc$status==0),2)
paste('Percent of cases sampled as control in Control group:',control_to_case)

paste('Number of unique cases sampled as control:',ramq_cc%>%filter(tcase==1&status==0)%>%summarise(n=n_distinct(id)))
#mean length of follow-up for those controls:
a<-semi_join(ramq_cc[,c('id','tfu')],ramq_cc%>%filter(tcase==1&status==0)%>%
                        distinct(id,tfu))
paste('The Average follow-up time for those cases in the Control group:',round(a%>%summarise(mean=mean(tfu)),2))
```

From the result, about `r control_to_case` percent of the controls became cases at a later time point after being sampled as control. Also, a high proportion of cases were sampled as control. Therefore, there is a high degree of similarity between case and control groups in this study.

#### Exposure pattern between case and control
```{r}
#load exposure dataset (subset of all prescription data where ahf code==481024)
exposure<-readRDS('../RData files/LTRA.RData')

#subset exposure data to individuals in ramq_cc cohort:
exposure<-exposure%>%
          left_join(distinct(ramq_cc,id,tentry_date,texit_date),by=c('nam'='id'))%>%
          filter(!is.na(tentry_date))%>%  #select individuals in the ramq_cc dataset
          distinct()

paste0('Number of individuals from exposure set with 0 quantite and duree:',nrow(exposure%>%filter(quantite==0 & duree==0)))

```

```{r}
#subset exposure data to include prescription history within follow-up period:
#exposure<-exposure%>%filter(dt_serv>=tentry_date)

#first dt_serv <texit_date
#exposure<-exposure%>%group_by(nam)%>%mutate(first=min(dt_serv))%>%ungroup()%>%filter(first<=texit_date)


exposure<-exposure%>%filter(dt_serv>=tentry_date & dt_serv<=texit_date,
                            quantite>0)



#exposure<-exposure%>%filter(dt_serv<=texit_date,
#                            quantite>0)
```

```{r}
#discontinuation definition: dt_serv(current) > dt_serv(before)+duree+15
#more than 15 days gap between the start of current prescription and the end of last prescription

#recency definition: the days between the end of last prescription and exit date
exposure<-exposure%>%group_by(nam)%>%
                     arrange(dt_serv)%>%
                     mutate(interval=dt_serv-lag(dt_serv))%>%
                     mutate(discontinue=ifelse(interval-lag(duree)>=15,1,0))%>%
                     mutate(recency=as.numeric(last(texit_date)-last(dt_serv))-last(duree))%>%
                     mutate(start_to_exit=as.numeric(texit_date-first(dt_serv)))%>%
                     ungroup()
#assign NA's in discontinue to 0 due to first observation in each group
exposure$discontinue[is.na(exposure$discontinue)]<-0

#check to see whether most exposure are continuous exposure?
#look at distribution of difference between total duration vs (last prescription-first prescription)
exposure_time<-exposure%>%group_by(nam)%>%
                          mutate(total_dur=sum(duree),
                                 total_time=as.numeric(max(dt_serv)-min(dt_serv)),
                                 discontinue=sum(discontinue))%>%
                                 ungroup()%>%
                                 distinct(nam,total_dur,total_time,discontinue,
                                          recency,start_to_exit)%>%
                                 mutate(discontinue=as.factor(ifelse(discontinue>0,1,0)))
#if total_time==0, one time user
```

```{r,warning=F}
#combine exposure summary table to ramq_cc
ramq_cc_exposure<-left_join(ramq_cc[,c('caseid','id','status')],exposure_time,by=c('id'='nam'))
ramq_cc_exposure$exposure<-as.factor(ifelse(is.na(ramq_cc_exposure$total_time),0,1))
tb<-table(ramq_cc_exposure$exposure,ramq_cc_exposure$status)
rownames(tb)<-c('Unexposed','Exposed')
colnames(tb)<-c('Control','Case')
tb
```
#####Distribution of cumulative exposure time between cases and controls
```{r,warning=F,fig.height=4,fig.width=6}
ggplot(ramq_cc_exposure,aes(x=total_dur,fill=status,color=status))+geom_histogram(binwidth=20,alpha=0.3)+
  xlab('cumulative duration of exposure during follow-up (Days)')
tapply(ramq_cc_exposure$total_dur,ramq_cc_exposure$status,summary)
```

There is a fairly large proportion of individuals having very short cumulative exposure.
Is that because they start drug close to exit_date?

##### Distribution of first prescription to exit date in cases and controls:
```{r,warning=F,fig.height=4,fig.width=8}
#start to exit :defined as texit-first prescription
#total_time: defined as last(dt_serv)-first(dtserv)
#overlaying both graph to see if there are many individuals who continued prescription after case definition:
ggplot(data=ramq_cc_exposure,aes(x=start_to_exit,fill=status,color=status))+geom_histogram(binwidth=30,alpha=0.3)+
  xlab('Start of first prescription to exit_date (Days)')+ggtitle('Whole case-control cohort')

ggplot(data=exposure_time,aes(x=start_to_exit))+geom_histogram(bins=25,alpha=0.2)+ggtitle('Exposed individuals')+scale_x_continuous(breaks=c(0,1000,1500))
 
tapply(ramq_cc_exposure$start_to_exit,ramq_cc_exposure$status,summary)
```


#### Distribution of time from end of last prescription to exit date in cases and controls:
```{r,warning=F,fig.height=4,fig.width=6}
ggplot(ramq_cc_exposure,aes(x=recency,fill=status,color=status))+geom_histogram(bins=25,alpha=0.3)+
  xlab('End of last prescription to exit_date (Days)')+scale_y_continuous(limits=c(0,1200))
```

A majority of cases either were still taking the drug at the time becoming a case or stopped taking the drug within short period of time before.

##### Distribution of first and last year of exposure:
```{r}
#confirm with year of first prescription: (from exposure data)
exposure_date<-exposure%>%group_by(nam)%>%mutate(first=min(dt_serv),
                                                 last=max(dt_serv))%>%
                                          ungroup()%>%
                                          distinct(nam,first,last)

```

```{r,fig.height=4,fig.width=10}
p1<-ggplot(data=exposure_date,aes(x=year(first)))+geom_bar()+ggtitle('First LTRA prescription by year')
p2<-ggplot(data=exposure_date,aes(x=year(last)))+geom_bar()+ggtitle('Last LTRA prescription by year')
gridExtra::grid.arrange(p1,p2,ncol=2,nrow=1)
```


Number of one-time user in cases and controls:
One-time user here was defined by the number of available dt_serv equals to 1
Some of the counts can come from patients who just start the prescription before exit the study.

```{r}
paste0('In Cases:',sum(ramq_cc_exposure$total_time==0&ramq_cc_exposure$status==1,na.rm=T))
paste0('In Controls:',sum(ramq_cc_exposure$total_time==0&ramq_cc_exposure$status==0,na.rm=T))

```

Number of individuals who had discontinuation of prescription during follow-up:
```{r}
tb2<-table(ramq_cc_exposure$discontinue,ramq_cc_exposure$status)
rownames(tb2)<-c('No discontinuation','Discontinuation')
colnames(tb2)<-c('Control','Case')
rbind(tb2,round(prop.table(tb2,2)[2,],2))
```

#### Apply cut-off of 90 days to exposure:
```{r}
ramq_cc_exposure$exposure<-ifelse(ramq_cc_exposure$total_dur>=90,1,0)
#convert NA's to 0:
ramq_cc_exposure$exposure[is.na(ramq_cc_exposure$exposure)]<-0
ramq_cc_exposure$exposure<-as.factor(ramq_cc_exposure$exposure)

tb3<-table(ramq_cc_exposure$exposure,ramq_cc_exposure$status)
colnames(tb3)<-c('Control','Case')
rownames(tb3)<-c('No exposure','Exposure')
tb3
round(prop.table(tb3,2)[2,],4)
```
```{r}
#define exposure 2 with current user or not:
#current user defined as the gap between the end of last prescription and cohort exit date is less
#than 7 days
ramq_cc_exposure$exposure2<-ifelse(ramq_cc_exposure$recency<=7,1,0)
#convert NA's to 0:
ramq_cc_exposure$exposure2[is.na(ramq_cc_exposure$exposure2)]<-0
ramq_cc_exposure$exposure2<-as.factor(ramq_cc_exposure$exposure2)

tb4<-table(ramq_cc_exposure$exposure2,ramq_cc_exposure$status)
colnames(tb4)<-c('Control','Case')
rownames(tb4)<-c('No exposure','Exposure')
tb4
round(prop.table(tb4,2)[2,],4)
```

#### Comorbidity statistics:{.tabset}
##### Overall combined {-}
```{r}
variables<-c('tageind','tfu','sexe','charlson','diabete_combined','ckd_combined','hypertension_combined',
             'pvd_combined','cad_combined','copd_combined','hyperlipidemia_combined','hf_combined')
factor_vars<-c('sexe','diabete_combined','ckd_combined','hypertension_combined',
               'pvd_combined','cad_combined','copd_combined','hyperlipidemia_combined','hf_combined')

CreateTableOne(variables,'status',ramq_cc,factorVars =factor_vars )
```

##### Diagnostic and Prescription {-}
```{r}
variables<-c('tageind','tfu','sexe','dyslipidemia','diabete','CKD','hypertension','COPD','CAD','peripheral','HF',
             'theophylline','beta_agonist','inhaled_corti','ACE_ARB','diuretic','alpha_blocker','calcium_channel','thiazolidinediones','sulfonylure',
             'biguanide','insulin','statin','nasaid','beta_blocker')
factor_vars<-c('sexe','dyslipidemia','diabete','CKD','hypertension','COPD','CAD','peripheral','HF','theophylline',
               'beta_agonist','inhaled_corti','ACE_ARB','diuretic','alpha_blocker','calcium_channel',
               'thiazolidinediones','sulfonylure','biguanide','insulin','statin','nasaid','beta_blocker')

CreateTableOne(variables,'status',ramq_cc,factorVars =factor_vars )
```

#### Conditional logistic model
```{r}
ramq_cc$status<-as.numeric(ramq_cc$status)
ramq_cc$pair<-rleid(ramq_cc$caseid)
ramq_cc$exposure<-ramq_cc_exposure$exposure
model<-clogit(data=ramq_cc,status~exposure+sexe+diabete_combined+ckd_combined+hypertension_combined+pvd_combined+copd_combined+hyperlipidemia_combined+hf_combined+statin+strata(pair))

summary(model)


#calculate unadjusted OR:
model_unadj<-clogit(data=ramq_cc,status~exposure+strata(pair))
summary(model_unadj)
```

```{r,eval=F}
#test for colinearity with VIF test from caret package:
library(car)



#model selection:
model.null<-clogit(data=ramq_cc,status~exposure+strata(pair))
model.full<-clogit(data=ramq_cc,status~exposure+sexe+diabete_combined+ckd_combined+hypertension_combined+pvd_combined+copd_combined+hyperlipidemia_combined+hf_combined+charlson+statin+strata(pair))

library(MASS)
#backward selection AIC:
backward<-stepAIC(model.full,direction='backward',trace=0)
summary(backward)

#forward selection AIC
forward<-step(model.null,scope=list(lower=model.null,upper=model.full),
               direction='forward',trace=0)

summary(forward)

vif(forward)


```

After model selection: hypertension, pvd, hyperlipidemia were dropped from the model

With Reduced model: the adjusted OR is 1.21 with 95% CI (1.10,1.33), not changed much from full model


#### Other models
Also we want to investigate the OR between current LTRAs user vs past LTRAs user
The current users are defined as individuals whose last LTRA prescription ends 7 days or less before cohort exit date.

```{r}
ramq_cc$exposure2<-ramq_cc_exposure$exposure2
model2<-clogit(data=ramq_cc,status~exposure2+sexe+diabete_combined+ckd_combined+hypertension_combined+pvd_combined+copd_combined+hyperlipidemia_combined+hf_combined+charlson+strata(pair))

summary(model2)
```

```{r}
model2_unadj<-clogit(data=ramq_cc,status~exposure2+strata(pair))
summary(model2_unadj)
```

The OR become non-significant, probably due to the elimination of 90 days cut-off.
Now let's define exposure as individuals who are both current users and have at least 90 day coverage.

```{r}
ramq_cc$exposure3<-as.factor(ifelse(ramq_cc$exposure==1 & ramq_cc$exposure2==1,1,0))

table(ramq_cc$exposure3,ramq_cc$status)
model3<-clogit(data=ramq_cc,status~exposure3+sexe+diabete_combined+ckd_combined+hypertension_combined+pvd_combined+copd_combined+hyperlipidemia_combined+hf_combined+charlson+statin+strata(pair))

summary(model3)
```

```{r}
model3_unadj<-clogit(data=ramq_cc,status~exposure3+strata(pair))
summary(model3_unadj)
```

